fdcoexist

### 2018-08-17 paper_simulations.Rmd
:sparkles: :hammer: Implement high correlations simul + modularize fcts for figures

---
title: "Full Simulations for the paper"
author:
    - Pierre Denelle
    - Matthias Grenié
    - Cyrille Violle
    - Caroline M. Tucker
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

@@

tidy_simulation = function(scenario, given_simul, given_traits,
                           given_trait_scenar) {
    landscape_df(given_simul[[scenario]], given_traits,
                 trait_scenarios[[scenario]]) %>%
        inner_join(given_traits %>%
                       as.data.frame() %>%
                       rownames_to_column("species"),
                   by = "species") %>%
        group_by(patch) %>%
        mutate(rel_abund = N/sum(N)) %>%
        ungroup()
}
# Initial parameters -----------------------------------------------------------
devtools::load_all()

@@

facet_scenar = facet_wrap(vars(scenario), labeller = as_labeller(trait_scenar_names))
```

@@

```{r no_compet_no_cor_abund_env}
plot_no_compet_no_cor_abund_env = ggplot(no_compet_no_cor_df %>%
           filter(N >= 2), aes(patch_optim, N)) +
    geom_point(shape = ".") +
    geom_smooth() +
    facet_scenar +
    labs(x = "Environmental value",
         y = "Species Abundance (gen. = 150)",
         subtitle = no_compet_no_cor_label)

@@

```{r no_compet_no_cor_cwm_env}
plot_no_compet_no_cor_cwm_env = no_compet_no_cor_df %>%
    filter(N >= 2) %>%
    group_by(scenario, patch) %>%
    summarise_at(vars(contains("cwm")), sum, na.rm = TRUE) %>%
    mutate(patch_optim = gsub("patches", "", patch) %>%
               as.numeric()) %>%
    gather("cwm_name", "cwm_value", contains("cwm")) %>%
    ggplot(aes(patch_optim, cwm_value, color = cwm_name)) +
    geom_point() +
    geom_smooth() +
    facet_scenar +
    scale_color_discrete(labels = c(trait1_cwm = "Trait 1",
                                    trait2_cwm = "Trait 2",
                                    trait3_cwm = "Trait 3")) +
    labs(x = "Environment",
         y = "Community-Weighted Mean (CWM)",
         color = NULL,
         subtitle = no_compet_no_cor_label) +
    theme(legend.position = "top")

plot_no_compet_no_cor_cwm_env
```

```{r no_compet_no_cor_cwm_env_abund}
plot_no_compet_no_cor_cwm_env_abund = no_compet_no_cor_df %>%
    group_by(scenario, patch) %>%
    summarise_at(vars(contains("cwm"), N), sum) %>%
    mutate(patch_optim = gsub("patches", "", patch) %>%
               as.numeric()) %>%
    gather("cwm_name", "cwm_value", contains("cwm")) %>%
    ungroup() %>%
    ggplot(aes(patch_optim, cwm_value, z = N)) +
    stat_summary_2d(bins = 24) +
    facet_grid(vars(scenario), vars(cwm_name),
               labeller = labeller(scenario = trait_scenar_names,
                                   cwm_name = label_value)) +
    scale_fill_viridis_c(trans = "log10") +
    labs(x = "Environment",
         y = "Community-Weighted Mean (CWM)",
         fill = "Total Abundance",
         subtitle = no_compet_no_cor_label) +
    theme(legend.position = "top")

@@

```{r no_compet_no_cor_df_ind_landscape}
plot_no_compet_no_cor_df_ind_landscape = no_compet_no_cor_df %>%
    select(-contains("cwm")) %>%
    gather("trait_name", "trait_value", contains("trait")) %>%
    ggplot(aes(patch_optim, trait_value, z = N)) +
    stat_summary_2d(bins = 22) +
    facet_grid(vars(scenario), vars(trait_name),
               labeller = labeller(scenario = trait_scenar_names,
                                   trait_name = function(x) gsub("_", " ", x))) +
    scale_fill_viridis_c() +
    labs(x = "Environment",
         y = "Trait Value",
         fill = "Species Abundance",
         subtitle = no_compet_no_cor_label) +
    theme(legend.position = "top")

@@

```{r no_compet_low_cor_abund_env}
plot_no_compet_low_cor_abund_env = ggplot(no_compet_low_cor_df %>%
           filter(N >= 2), aes(patch_optim, N)) +
    geom_point(shape = ".") +
    geom_smooth() +
    facet_scenar +
    labs(x = "Environmental value",
         y = "Species Abundance (gen. = 150)",
         subtitle = no_compet_low_cor_label)

@@

```{r no_compet_low_cor_cwm_env}
plot_no_compet_low_cor_cwm_env = no_compet_low_cor_df %>%
    filter(N >= 2) %>%
    group_by(scenario, patch) %>%
    summarise_at(vars(contains("cwm")), sum, na.rm = TRUE) %>%
    mutate(patch_optim = gsub("patches", "", patch) %>%
               as.numeric()) %>%
    gather("cwm_name", "cwm_value", contains("cwm")) %>%
    ggplot(aes(patch_optim, cwm_value, color = cwm_name)) +
    geom_point() +
    geom_smooth() +
    facet_scenar +
    scale_color_discrete(labels = c(trait1_cwm = "Trait 1",
                                    trait2_cwm = "Trait 2",
                                    trait3_cwm = "Trait 3")) +
    labs(x = "Environment",
         y = "Community-Weighted Mean (CWM)",
         color = NULL,
         subtitle = no_compet_low_cor_label) +
    theme(legend.position = "top")

@@

```{r no_compet_low_cor_cwm_env_abund}
plot_no_compet_low_cor_cwm_env_abund = no_compet_low_cor_df %>%
    group_by(scenario, patch) %>%
    summarise_at(vars(contains("cwm"), N), sum) %>%
    mutate(patch_optim = gsub("patches", "", patch) %>%
               as.numeric()) %>%
    gather("cwm_name", "cwm_value", contains("cwm")) %>%
    ungroup() %>%
    ggplot(aes(patch_optim, cwm_value, z = N)) +
    stat_summary_2d(bins = 24) +
    facet_grid(vars(scenario), vars(cwm_name),
               labeller = labeller(scenario = trait_scenar_names,
                                   cwm_name = label_value)) +
    scale_fill_viridis_c(trans = "log10") +
    labs(x = "Environment",
         y = "Community-Weighted Mean (CWM)",
         fill = "Total Abundance",
         subtitle = no_compet_low_cor_label) +
    theme(legend.position = "top")

@@

```{r no_compet_low_cor_df_ind_landscape}
plot_no_compet_low_cor_df_ind_landscape = no_compet_low_cor_df %>%
    select(-contains("cwm")) %>%
    gather("trait_name", "trait_value", contains("trait")) %>%
    ggplot(aes(patch_optim, trait_value, z = N)) +
    stat_summary_2d(bins = 17) +
    facet_grid(vars(scenario), vars(trait_name),
               labeller = labeller(scenario = trait_scenar_names,
                                   trait_name = function(x) gsub("_", " ", x))) +
    scale_fill_viridis_c() +
    labs(x = "Environment",
         y = "Trait Value",
         fill = "Species Abundance",
         subtitle = no_compet_low_cor_label) +
    theme(legend.position = "top")

@@

no_compet_high_cor_label = paste0("No competition; ",
                                d * 100, "% dispersal;",
                                " 3 low correlated (r = 0.7) traits")
```

```{r no_compet_high_cor_abund_env}
plot_no_compet_high_cor_abund_env = ggplot(no_compet_high_cor_df %>%
           filter(N >= 2), aes(patch_optim, N)) +
    geom_point(shape = ".") +
    geom_smooth() +
    facet_scenar +
    labs(x = "Environmental value",
         y = "Species Abundance (gen. = 150)",
         subtitle = no_compet_high_cor_label)

@@

```{r no_compet_high_cor_cwm_env}
plot_no_compet_high_cor_cwm_env = no_compet_high_cor_df %>%
    filter(N >= 2) %>%
    group_by(scenario, patch) %>%
    summarise_at(vars(contains("cwm")), sum, na.rm = TRUE) %>%
    mutate(patch_optim = gsub("patches", "", patch) %>%
               as.numeric()) %>%
    gather("cwm_name", "cwm_value", contains("cwm")) %>%
    ggplot(aes(patch_optim, cwm_value, color = cwm_name)) +
    geom_point() +
    geom_smooth() +
    facet_scenar +
    scale_color_discrete(labels = c(trait1_cwm = "Trait 1",
                                    trait2_cwm = "Trait 2",
                                    trait3_cwm = "Trait 3")) +
    labs(x = "Environment",
         y = "Community-Weighted Mean (CWM)",
         color = NULL,
         subtitle = no_compet_high_cor_label) +
    theme(legend.position = "top")

plot_no_compet_high_cor_cwm_env
```

```{r no_compet_high_cor_cwm_env_abund}
plot_no_compet_high_cor_cwm_env_abund = no_compet_high_cor_df %>%
    group_by(scenario, patch) %>%
    summarise_at(vars(contains("cwm"), N), sum) %>%
    mutate(patch_optim = gsub("patches", "", patch) %>%
               as.numeric()) %>%
    gather("cwm_name", "cwm_value", contains("cwm")) %>%
    ungroup() %>%
    ggplot(aes(patch_optim, cwm_value, z = N)) +
    stat_summary_2d(bins = 24) +
    facet_grid(vars(scenario), vars(cwm_name),
               labeller = labeller(scenario = trait_scenar_names,
                                   cwm_name = label_value)) +
    scale_fill_viridis_c(trans = "log10") +
    labs(x = "Environment",
         y = "Community-Weighted Mean (CWM)",
         fill = "Total Abundance",
         subtitle = no_compet_high_cor_label) +
    theme(legend.position = "top")

@@

```{r no_compet_high_cor_df_ind_landscape}
plot_no_compet_high_cor_df_ind_landscape = no_compet_high_cor_df %>%
    select(-contains("cwm")) %>%
    gather("trait_name", "trait_value", contains("trait")) %>%
    ggplot(aes(patch_optim, trait_value, z = N)) +
    stat_summary_2d(bins = 17) +
    facet_grid(vars(scenario), vars(trait_name),
               labeller = labeller(scenario = trait_scenar_names,
                                   trait_name = function(x) gsub("_", " ", x))) +
    scale_fill_viridis_c() +
    labs(x = "Environment",
         y = "Trait Value",
         fill = "Species Abundance",
         subtitle = no_compet_high_cor_label) +
    theme(legend.position = "top")

@@

@@

@@

>>>>

---
title: "fdcoexist"
author:
    - Pierre Denelle, Matthias Grenié, Cyrille Violle and Caroline M. Tucker
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

@@

tidy_simulation = function(scenario, given_simul, given_traits,
                           given_trait_scenar) {
    landscape_df(given_simul[[scenario]], given_traits,
                 trait_scenarios[[scenario]]) %>%
        inner_join(given_traits %>%
                       as.data.frame() %>%
                       rownames_to_column("species"),
                   by = "species") %>%
        group_by(patch) %>%
        mutate(rel_abund = N/sum(N)) %>%
        ungroup()
}
# Initial parameters -----------------------------------------------------------
suppressMessages(devtools::load_all())

@@

# Graphical functions ----------------------------------------------------------
facet_scenar = facet_wrap(vars(scenario),
                          labeller = as_labeller(trait_scenar_names))

# Plot the relationship between species abundances along the environment in the
# different scenarios
fig_abund_env = function(simul_df, simul_label) {
    ggplot(simul_df %>%
           filter(N >= 2), aes(patch_optim, N)) +
    geom_point(shape = ".") +
    geom_smooth() +
    facet_scenar +
    labs(x = "Environmental value",
         y = "Species Abundance (gen. = 150)",
         subtitle = simul_label)
}

# Plot the relationship between the CWM of the different trait along the
# environment for the different scenarios
fig_cwm_env = function(simul_df, simul_label) {
    simul_df %>%
    filter(N >= 2) %>%
    group_by(scenario, patch) %>%
    summarise_at(vars(contains("cwm")), sum, na.rm = TRUE) %>%
    mutate(patch_optim = gsub("patches", "", patch) %>%
               as.numeric()) %>%
    gather("cwm_name", "cwm_value", contains("cwm")) %>%
    ggplot(aes(patch_optim, cwm_value, color = cwm_name)) +
    geom_point() +
    geom_smooth() +
    facet_scenar +
    scale_color_discrete(labels = c(trait1_cwm = "Trait 1",
                                    trait2_cwm = "Trait 2",
                                    trait3_cwm = "Trait 3")) +
    labs(x = "Environment",
         y = "Community-Weighted Mean (CWM)",
         color = NULL,
         subtitle = simul_label) +
    theme(legend.position = "top")
}

# Plot the relationship between patch CWMs along the environment considering
# total abundance
fig_cwm_env_abund = function(simul_df, simul_label, n_bins = 24,
                             given_trans = "identity") {
    simul_df %>%
    group_by(scenario, patch) %>%
    summarise_at(vars(contains("cwm"), N), sum) %>%
    mutate(patch_optim = gsub("patches", "", patch) %>%
               as.numeric()) %>%
    gather("cwm_name", "cwm_value", contains("cwm")) %>%
    ungroup() %>%
    ggplot(aes(patch_optim, cwm_value, z = N)) +
    stat_summary_2d(bins = n_bins) +
    facet_grid(vars(scenario), vars(cwm_name),
               labeller = labeller(scenario = trait_scenar_names,
                                   cwm_name = label_value)) +
    scale_fill_viridis_c(trans = given_trans) +
    labs(x = "Environment",
         y = "Community-Weighted Mean (CWM)",
         fill = "Total Abundance",
         subtitle = simul_label) +
    theme(legend.position = "top")
}

# Relationship betwee, species trait values along the enviroment showing their
# local abundances for the different traits considered
fig_ind_landscape = function(simul_df, simul_label, n_bins = 22,
                             given_trans = "identity") {
    simul_df %>%
    select(-contains("cwm")) %>%
    gather("trait_name", "trait_value", contains("trait")) %>%
    ggplot(aes(patch_optim, trait_value, z = N)) +
    stat_summary_2d(geom = "raster", bins = n_bins, interpolate = TRUE) +
    facet_grid(vars(scenario), vars(trait_name),
               labeller = labeller(scenario = trait_scenar_names,
                                   trait_name = function(x) gsub("_", " ", x))) +
    scale_fill_viridis_c(trans = given_trans) +
    labs(x = "Environment",
         y = "Trait Value",
         fill = "Species Abundance",
         subtitle = simul_label) +
    theme(legend.position = "top")
}
```

@@

```{r no_compet_no_cor_abund_env}
plot_no_compet_no_cor_abund_env = fig_abund_env(no_compet_no_cor_df,
                                                no_compet_no_cor_label)

@@

```{r no_compet_no_cor_cwm_env}
plot_no_compet_no_cor_cwm_env = fig_cwm_env(no_compet_no_cor_df,
                                            no_compet_no_cor_label)
plot_no_compet_no_cor_cwm_env
```

```{r no_compet_no_cor_cwm_env_abund}
plot_no_compet_no_cor_cwm_env_abund = fig_cwm_env_abund(no_compet_no_cor_df,
                                                        no_compet_no_cor_label,
                                                        given_trans = "log10")

@@

```{r no_compet_no_cor_df_ind_landscape}
plot_no_compet_no_cor_df_ind_landscape = fig_ind_landscape(no_compet_no_cor_df,
                                                           no_compet_no_cor_label,
                                                           n_bins = 22,
                                                           given_trans = "identity")

@@

```{r no_compet_low_cor_abund_env}
plot_no_compet_low_cor_abund_env = fig_abund_env(no_compet_low_cor_df,
                                                 no_compet_low_cor_label)

@@

```{r no_compet_low_cor_cwm_env}
plot_no_compet_low_cor_cwm_env = fig_cwm_env(no_compet_low_cor_df,
                                             no_compet_low_cor_label)

@@

```{r no_compet_low_cor_cwm_env_abund}
plot_no_compet_low_cor_cwm_env_abund = fig_cwm_env_abund(no_compet_low_cor_df,
                                                         no_compet_low_cor_label,
                                                         given_trans = "log10")

@@

```{r no_compet_low_cor_df_ind_landscape}
plot_no_compet_low_cor_df_ind_landscape = fig_ind_landscape(no_compet_low_cor_df,
                                                            no_compet_low_cor_label,
                                                            n_bins = 17,
                                                            given_trans = "identity")

@@

no_compet_high_cor_label = paste0("No competition; ",
                                d * 100, "% dispersal;",
                                " 3 high correlated (r = 0.7) traits")
```

```{r no_compet_high_cor_abund_env}
plot_no_compet_high_cor_abund_env = fig_abund_env(no_compet_high_cor_df,
                                                  no_compet_high_cor_label)

@@

```{r no_compet_high_cor_cwm_env}
plot_no_compet_high_cor_cwm_env = fig_cwm_env(no_compet_high_cor_df,
                                              no_compet_high_cor_label)
plot_no_compet_high_cor_cwm_env
```

```{r no_compet_high_cor_cwm_env_abund}
plot_no_compet_high_cor_cwm_env_abund = fig_cwm_env_abund(no_compet_high_cor_df,
                                                          no_compet_high_cor_label,
                                                          given_trans = "log10")

@@

```{r no_compet_high_cor_df_ind_landscape}
plot_no_compet_high_cor_df_ind_landscape = fig_ind_landscape(no_compet_high_cor_df,
                                                             no_compet_high_cor_label,
                                                             n_bins = 17)

@@

```{r simul_compet_no_corr}
compet_no_cor <- lapply(
    trait_scenarios,
    function(x){
        multigen(traits = uncor_traits,
                 trait_weights = x,
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = A, d = d,
                 k = k, width = env_width)
    })

# Takes simulation results and put them as an exploitable data.frame with only
# gen = 150 abundances
compet_no_cor_df = tidy_simulation_list(compet_no_cor,
                                        set_of_traits = uncor_traits,
                                        trait_scenarios = trait_scenarios)

compet_no_cor_label = paste0("Competition (A = ", A, "); ",
                             d * 100, "% dispersal;",
                             " 3 uncorrelated traits")
```

```{r compet_no_cor_abund_env}
plot_compet_no_cor_abund_env = fig_abund_env(compet_no_cor_df,
                                             compet_no_cor_label)

plot_compet_no_cor_abund_env
```

```{r compet_no_cor_cwm_env}
plot_compet_no_cor_cwm_env = fig_cwm_env(compet_no_cor_df,
                                         compet_no_cor_label)

plot_compet_no_cor_cwm_env
```

```{r compet_no_cor_cwm_env_abund}
plot_compet_no_cor_cwm_env_abund = fig_cwm_env_abund(compet_no_cor_df,
                                                     compet_no_cor_label,
                                                     given_trans = "identity")
plot_compet_no_cor_cwm_env_abund
```

```{r compet_no_cor_df_ind_landscape}
plot_compet_no_cor_df_ind_landscape = fig_ind_landscape(compet_no_cor_df,
                                                        compet_no_cor_label,
                                                        n_bins = 22)

plot_compet_no_cor_df_ind_landscape
```

@@

```{r simul_compet_low_corr}
compet_low_cor <- lapply(
    trait_scenarios,
    function(x){
        multigen(traits = uncor_traits,
                 trait_weights = x,
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = A, d = d,
                 k = k, width = env_width)
    })

# Takes simulation results and put them as an exploitable data.frame with only
# gen = 150 abundances
compet_low_cor_df = tidy_simulation_list(compet_low_cor,
                                        set_of_traits = low_cor_traits,
                                        trait_scenarios = trait_scenarios)

compet_low_cor_label = paste0("Competition (A = ", A, "); ",
                              d * 100, "% dispersal;",
                              " 3 correlated traits (r = 0.3)")
```

```{r compet_low_cor_abund_env}
plot_compet_low_cor_abund_env = fig_abund_env(compet_low_cor_df,
                                              compet_low_cor_label)

plot_compet_low_cor_abund_env
```

```{r compet_low_cor_cwm_env}
plot_compet_low_cor_cwm_env = fig_cwm_env(compet_low_cor_df,
                                          compet_low_cor_label)

plot_compet_low_cor_cwm_env
```

```{r compet_low_cor_cwm_env_abund}
plot_compet_low_cor_cwm_env_abund = fig_cwm_env_abund(compet_low_cor_df,
                                                      compet_low_cor_label,
                                                      given_trans = "identity")

plot_compet_low_cor_cwm_env_abund
```

```{r compet_low_cor_df_ind_landscape}
plot_compet_low_cor_df_ind_landscape = fig_ind_landscape(compet_low_cor_df,
                                                         compet_low_cor_label,
                                                         n_bins = 17)

plot_compet_low_cor_df_ind_landscape
```

@@

```{r simul_compet_high_corr}
compet_high_cor <- lapply(
    trait_scenarios,
    function(x){
        multigen(traits = uncor_traits,
                 trait_weights = x,
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = A, d = d,
                 k = k, width = env_width)
    })

# Takes simulation results and put them as an exploitable data.frame with only
# gen = 150 abundances
compet_high_cor_df = tidy_simulation_list(compet_high_cor,
                                        set_of_traits = high_cor_traits,
                                        trait_scenarios = trait_scenarios)

compet_high_cor_label = paste0("Competition (A = ", A, "); ",
                             d * 100, "% dispersal;",
                             " 3 correlated traits (r = 0.7)")
```

```{r compet_high_cor_abund_env}
plot_compet_high_cor_abund_env = fig_abund_env(compet_high_cor_df,
                                               compet_high_cor_label)

plot_compet_high_cor_abund_env
```

```{r compet_high_cor_cwm_env}
plot_compet_high_cor_cwm_env = fig_cwm_env(compet_high_cor_df,
                                           compet_high_cor_label)

plot_compet_high_cor_cwm_env
```

```{r compet_high_cor_cwm_env_abund}
plot_compet_high_cor_cwm_env_abund = fig_cwm_env_abund(compet_high_cor_df,
                                                       compet_high_cor_label,
                                                       given_trans = "identity")

plot_compet_high_cor_cwm_env_abund
```

```{r compet_high_cor_df_ind_landscape}
plot_compet_high_cor_df_ind_landscape = fig_ind_landscape(compet_high_cor_df,
                                                          compet_high_cor_label,
                                                          n_bins = 17)
plot_compet_high_cor_df_ind_landscape
```
